Global Namespaces Scale as Well as Decoupled Namespaces

### 2017-01-31 paper/paper.tex
paper: finish related work
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}
Today's large and highly-parallel workloads are bottlenecked by metadata
services because many processes end up accessing the same shared resoure. In
HPC, state-of-the-art file systems are abandoning POSIX because the
sychronization and serializiation overheads are too costly -- and sometimes
even unnceccessary -- for some of their applications.  While the performance
benefits are plain for these users, other applications that rely on stronger
consistency must be re-written or deployed on a different system. We present
Cudele, a programmable file system that supports different degrees of
consistency and fault tolerance within the same namespace.  In this paper we
examine a general purpose file system and describe the overheads of the design
decisions that were made; namely strong consistency and fault tolerance. Then,
using Cudele, we relax those constraints on specific subtrees in the global
namespace and show improved performance without sacrificing the strong
consistency needs of other users.
\end{abstract}

@@

Decoupling the namespaces has many advantages, including improved scalability,
higher resource utilization, and better performance.
BatchFS~\cite{zheng:pdsw2014-batchfs} and DeltaFS~\cite{zheng:pdsw2015-deltafs}
are near-POSIX filesystems that give clients the ability to decouple subtrees
from the namespace so that the applications can execute metadata operations
without synchronization and serialization.  These operations are applied to a
local snapshot of the file system namespace and conflicts are resolved either
by the application or by an external service . Applications link into a
metadata server library to reduce resource utilization and code paths (e.g., no
daemons and less interprocess communication). Unfortunately, decoupling the
namespaces has costs: (1) merging metadata state back into the global namespace
is slow; (2) failures are local to the failing node; and (3) the systems are
not backwards compatible.

@@

For (2), if the client fails and stays down, all metadata operations on the
decoupled namespace are lost. If the client recovers, the on-disk structures
(for BatchFS and DeltaFS this is the SSTables used in
TableFS~\cite{ren:atc2013-tablefs}) can be recovered. In other words, the
clients have state that cannot be recovered if the node stays failed and any
progress made will be lost. This scenario is a disaster for checkpoint-restart
where missed cycles may cause the checkpoint to bleed over into computation
time.

@@

\begin{table}[t!]
\centering
\begin{tabular}{l|l|l|l}
                  & decoupled~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs}
                  & global~\cite{weil:sc2004-dyn-metadata, ren:sc2014-indexfs}
                  & implied~\cite{bent_plfs_2009} \\ \hline
  consistency     & eventual      & capabilities  & naming \\
  fault tol.      & local         & journal       & journal
\end{tabular}
\caption{Each technique for handling the namespace has its trade-offs.
Decoupled namespaces have the highest throughput because they have weaker
consistency and fault tolerance semantics.
\label{table:namespaces}}
\end{table}

@@

\subsection{Fault Tolerance}
\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-journal.png}
\caption{Performance improves with larger journal segments because the metadata
server spends less time flushing the journal }\label{fig:throughput-journal}
\end{figure*}

@@

% purpose of the journal
CephFS uses RADOS (1) as a metadata store for all information about files
including the hierarchical namespace and (2) as a staging area for the journal
of updates before they are applied to the metadata store. (1) is essentially a
cache that improves performance by serving metadata from memory and (2) is the
mechanisms for achieving fault tolerance.

@@

\section{notes}
Linking clients into our custom libcephfs

Use namespace's recursive data structure to put policies on subtrees
- consistency: eventual vs. strong, global vs. local
  - e.g., BatchFS/DeltaFS: eventual, local
  - e.g., POSIX: strong, global
  - e.g., PLFS: no consistency
- fault tolerance: global vs. local
  - e.g., CephFS: global
  - e.g., BatchFS/DeltaFS: local

Experimental Setup
- Ceph: 9 OSDs, 1 MDS, 2 kernel client
- Workload limitations: blah

Workload: creates

Baseline: 200K creates in the same directory
- throughput: degrades at 950s
- CPU utilization: more at 950s
- inode cache: eviction dominate
- inodes +- to cache: eviction dominate
- per-disk throughput: RADOS not bottleneck

Experiment 1: Interference

\section{Experiments}

@@

\subsection{Baseline}
Experiment 0: creates in the same directory
- setup: why we use caching, we use the kernel client, how we circumvent max fragment size

Experiment 0: creates with a stat
- Hypothesis: metadata read pauses creates and requires a snapshot in time
  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
- can we delay snapshot?

Experiment 1: creates with a readdir
- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
- client0: create 100k, client1: stat at 2 mins

Experiment 2: scale the number of files
- See if the open/close spike occurs 
- Try to see why open/close spike is allowed to happen
- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
- client0: create 100k, client1: touch at 2 mins

Experiment 3: see how fast the cache satisfies a read
- client0: create 100k, stat inodes
- client0: create 100k, client1: stat inodes

lient 0: creates, client 1 create(s)

>>>>

% I work with the file system layer that sits on top of Ceph
% - internal subsystems and interfaces that Ceph uses to provide a strongly consistent POSIX file system
%   - e.g., load balancing mechanisms, for hotspots/flash crowds on directories
%     caches for inodes, capabilities for guarding a shared resource, and a large journal for fault tolerance
% - our programmability initiative talks about composing these internal interfaces for higher level services

% 
% Previous work: 
% - master's thesis: scale-out vs. scale-up architectures 
% - work in industry: file systems for the cloud, big data processing over
% object storage, and automating workflows
%
% I work under Carlos at UC Santa Cruz and we study how to migrate file system
% metadata to alleviate overloaded servers. The idea is to take the namespace
% on one metadata server, chop it up into smaller subtrees, and farm those
% subtrees out to other metadata servers.  It's very similar to IndexFS/ShardFS
% but we try to leverage namespace locality by allowing variable sized
% subtrees. We published a paper at SC'15 where we talk about a load balancing
% API we built on CephFS that let's administrators control how to chop up
% subtrees and when/where to move them (also got merged into Ceph).
%
% Go to conclusion
%
% Why I want to work with Brad
% - new perspectives on file system designs, especially NON-POSIXy 
% - work in a national lab: I know about theirarchitectures and problems but
% I've never used an HPC system - real world applications and benchmarks; most
% of my work has been accepted for novelty but not for fully benchmarking a
% system
% - lines up with my thesis
%
% What I want to work on:
% - continue working on my thesis; this next phase of having a global namespace
% but different consistency semantics within that namespace for data management
%   - namespace: recursive data structure as a logical model for understanding the user's goals
%   - merge: what is the proper way to merge deltafs/batchfs updates back into the global namespace
%   - data management: deltaFS made the case that apps shouldn't use the fs to synchronize;
%     we contend that you could use the fs as data management; remembering how to access your data
%
% What I have been working on 
% - leverage internal services/subsystems in Ceph; avoids code duplication and
% let's us test this over the same system 
% - benchmarking the overheads of strong consistency/journalling - we've figured
% out how to thrash capabilities and turn off the journal
\begin{abstract}

Today's large and highly-parallel workloads are bottlenecked by metadata
services because many processes end up accessing the same shared resoure. In
HPC, state-of-the-art file systems are abandoning POSIX because the
sychronization and serializiation overheads are too costly -- and sometimes
even unnceccessary -- for some of their applications.  While the performance
benefits are plain for these users, other applications that rely on stronger
consistency must be re-written or deployed on a different system.  We present
Cudele, a programmable file system that supports different degrees of
consistency and fault tolerance within the same namespace.  First, we take a
POSIX compliant file system and relax the consistency constraints by
implementing two metadata designs: delayed metadata update merge into the
global namespace and client local views of metadata updates.  Second, we store
the consistency and fault tolerance semantics in inodes so that subtrees within
the same namespace can be optimized for different workloads; the inodes are
programmable so that clients understand how to access the metadata in a
subtree.  Third, we present a theoretical framework for a metadata service and
a working implementation that provides the lowest common denominator needed to
implement a wide range of consistency/fault tolerance semantics that can be
benchmarked all on the same system.

\end{abstract}

% Questions
% - do you still collaborate with CMU? Would you want all of us to work together?
% - what projects do you want to work on?
% - how closely would you want to work with Carlos?
% - what system would you want to work on?

% The main difference between CephFS and the CMU systems is that our approach
% is a little more flexible.  We can partition the namespace in any way and can
% co-locate items together if they are dependent on each other -- in theroy...
% in practice it is very hard to figure out what the workload is trying to do.
% Also, we can program the decision making based on different metrics, like
% CPU, request rate, etc.

@@

% General
The bottlenecks associated with accessing POSIX file system metadata are not limited
to HPC workloads and the same challenges that plagued these systems for years are
finding their way into the cloud. Workloads that deal with many small files
({\it e.g.}, log processing and database
queries~\cite{thusoo:sigmod2010-facebook-infrastructure}) and large numbers of
simultaneous clients ({\it e.g.}, MapReduce
jobs~\cite{mckusick:acm2010-gfs-evolution}), are subject to the scalability of
the metadata service. The biggest challenge is that whenever a file
is touched the client must access the file's metadata and maintaining a file
system namespace imposes small, frequent accesses on the underlying storage
system~\cite{roselli:atec2000-FS-workloads}.  Unfortunately, scaling file
system metadata is a well-known problem and solutions for scaling data IO do
not work for metadata IO~\cite{roselli:atec2000-FS-workloads,
abad:techreport2012-fstrace, abad:ucc2012-mimesis,
alam:pdsw2011-metadata-scaling, weil:osdi2006-ceph}. There are two approaches
for improving the performance of metadata access.

\subsection{Metadata Load Balancing}

% approaches to load balancing (strong = more resuorces per unit work, weak =
% fixed resource per work unit)
One approach for improving metadata performance and scalability is to alleviate
overloaded servers by load balancing metadata IO across a cluster. Common
techniques include partitioning metadata when there are many writes and
replicating metadata when there are many reads. For example, IndexFS partitions
directories and clients write to different partitions by grabbing leases and
caching ancestor metadata for path traversal; it does well for strong scaling
because servers can keep more inodes in the cache which results in less RPCs.
Alternatively, ShardFS replicates directory state so servers do not need to
contact peers for path traversal; it does well for read workloads because all
file operations only require 1 RPC and for weak scaling because requests will
never incurr extra RPCs due to a full cache.  CephFS employs both techniques to
a lesser extent; directories can be replicated or sharded but the caching and
replication policies do not change depending on the balancing technique.
Despite the performance benefits these techniques add complexity and jeopardize
the robustness and performance characteristics of the metadata service because
the systems now need (1) policies to guide the migration decisions and (2)
mechanisms to address inconsistent states across servers.

% policies
Setting policies for migrations is arguably more difficult than adding the
migration mechanisms themselves.  For example, IndexFS and CephFS use the
GIGA+~\cite{GIGA+} technique for partitioning directories at a predefined
threshold and using lazy synchronization to redirect queries to the server that
``owns" the targeted metadata.  Determining when to partition directories and
when to migrate the directory fragments are policies that vary between systems:
GIGA+ partitions directories when the size reaches a certain number of files
and migrates directory fragments immediately; CephFS partitions directories
when they reach a threshold size or when the write temperature reaches a
certain value and migrates directory fragments when the hosting server has more
load than the other servers in the metadata cluster. Another policy is when and
how to replicate directory state; ShardFS replicates immediately and
pessimistically while CephFS replicates only when the read temperature reaches
a threshold.  There is a wide range of policies and it is difficult to address
with tunables and hard-coded design decisions.

% addressing inconsistency
In addition to the policies, distributing metadata across a cluster requiress
distributed transactions and cache coherence protocols to ensure strong
consistency ({e.g.}, POSIX).  For example, ShardFS pessimistically replicates
directory state and uses optimistic concurrency control for conflicts; namely
it does the operation and if there is a conflict at verification time it falls
back to two-phase locking.  Another example is IndexFS's inode cache which
reduces RPCs by caching ancestor paths -- the locality of this cache can be
thrashed by random reads but performs well for metadata writes. For
consistency, writes to directories in IndexFS block until the lease expires
while writes to direcotries in ShardFS are slow for everyone as it either
requires serialization or locking with many servers; reads in IndexFS are
subject to cache locality while reads in ShardFS always resolve to 1 RPC.
Another example of the overheads of addressing inconsistency is how CephFS
maintains client sessions and inode caches for capabilities (which in turn make
metadata access faster). When metadata is exchanged between metadata servers
these sessions/caches must be flushed and new statistics exchanged with a
scatter-gather process; this halts updates on the directories and blocks until
the authoratitive metadata server responds.  These protocols are discusssed in
more detail in the next section but their inclusion here is a testament to the
complexity of migrating metadata.

%For metadata writes it does a distributed transaction; monotonic writes with
%concurrent clients fail and do pessimistic locking through a mediated lock
%server to ensure strong consistency, non-monotonic writes grab locks at every
%server. Zooming in on monotonic writes: if permissions increase it executes on
%a primary then non-primary, if permissions decrease it executes on all
%non-primary then on primary.

The conclusion we have drawn from this related work is that metadata protocols
have a bigger impact on performance and scalability than load balancing.  
Understanding these protocols helps load balancing and gives us a better
understanding of the metrics we should use to make migration decisions ({\it
e.g.}, which operations reflect the state of the system), what types of
requests cause the most load, and how an overloaded system reacts ({\it e.g.},
increasing latencies, lower throughput, etc.).

\subsection{Relaxing POSIX}

% POSIX 
POSIX workloads require strong consistency and many file systems improve
performance by reducing the number of remote calls per operation ({\it i.e.}
RPC amplification). As discussed in the previous section, caching with leases and
replication are popular approaches to reducing the overheads of path traversals
but their performance is subject to cache locality and the amount of available
resources, respectively; for random workloads larger than the cache extra RPCs
hurt performance~\cite{IndexFS, CephFS} and for write heavy workloads with more
resources the RPCs for invalidations are harmful. Another approach to reducing
RPCs is to use leases or capabilities.

%IndexFS aggressively caches paths and
%handles permissions by handing out leases for metadata writes; metadata may
%only be modified when all leases have expired. 
%IndexFS~\cite{ren:sc2014-indexfs} aggressively caches pathnames and their
%permissions on the client servers. Modifications to metadata cached by clients
%is delayed until all client leases have expired. This reduces the RPC
%amplification to 1 when mutatating directory metadata ({\it e.g.,}
%\texttt{mkdir}, \texttt{chmod}, etc.) because clients are not querying MDSs for
%path traversals. The disadvantage of this approach is the high latency of
%mutation operations (reads to filenames).  ShardFS~\cite{xiao:socc2015-shardfs}
%replicates metadata (specifically directory lookup state) across the MDS
%cluster, reducing the RPC amplification to 1 for file operations ({e.g.,}
%\texttt{stat}, \texttt{chmod}, \texttt{chown}, etc.). Modifications to the
%directory lookup state are done with optimistic concurrecny control and fall
%back to retry if verification fails. The disadvantage of this appraoch is the
%high number of RPCs for maintaining directory metadata mutations (writes to
%directories).  CephFS also maintains an inode cache but its notion of leases
%are much shorter, on the order of microseconds.

% Non POSIX
High performance computing has unique requirements for file systems ({e.g.},
fast creates) and well-defined workloads (e.g., workflows) that make relaxing
POSIX sensible.  One popular approach is to allow clients to ``lock'' parts of
the namespace to improve performance and scalability by avoiding
synchronization, false sharing, and serialization.  BatchFS assumes the
application coordinates accesses to the namespace, so the clients can batch
local operations and merge with a global namespace image lazily. Similarily,
DeltaFS eliminates RPC traffic using subtree snapshots for non-conflicting
workloads and middleware for conflicting workloads. MarFS gives administrators
the ability to (1) lock ``project directories" and (2) allocate GPFS clusters
for demanding directory workloads. TwoTiers eliminates high-latencies by
storing metadata in a flash tier; apps lock the namespace so that metadata can
be accessed more quickly.  Unfortunately, decoupling the namespaces has costs:
(1) merging metadata state back into the global namespace is slow; (2) failures
are local to the failing node; and (3) the systems are not backwards
compatible.

% NON POSIX
%Decoupling the namespaces has many advantages, including improved scalability,
%higher resource utilization, and better performance.  BatchFS and DeltaFS are
%near-POSIX filesystems that give clients the ability to decouple subtrees from
%the namespace so that the applications can execute metadata operations without
%synchronization and serialization.  These operations are applied to a local
%snapshot of the file system namespace and conflicts are resolved either by the
%application or by an external service . Applications link into a metadata
%server library to reduce resource utilization and code paths (e.g., no daemons
%and less interprocess communication).

@@

For (2), if the client fails and stays down, all metadata operations on the
decoupled namespace are lost. If the client recovers, the on-disk structures
(for BatchFS and DeltaFS this is the SSTables used in TableFS) can be
recovered. In other words, the clients have state that cannot be recovered if
the node stays failed and any progress will be lost. This scenario is a
disaster for checkpoint-restart where missed cycles may cause the checkpoint to
bleed over into computation time.

@@

@@

\subsection{Fault Tolerance}
\begin{figure*}[tb] \centering
\includegraphics[width=1\textwidth]{./figures/journal.png} 
\caption{CephFS uses a journal to stage updates and tracks dirty metadata in
the collective memory of the MDSs. Each MDS maintains its own journal, which is
broken up into 4MB segments. These segments are pushed into RADOS and deleted
when that particular segment is trimmed from the end of the log. In addition to
journal segments, RADOS also stores per-directory objects. \label{fig:journal}}
\end{figure*}

\begin{figure*}[tb]%h
\centering
\includegraphics[width=180mm]{figures/throughput-journal.png}
\caption{Performance improves with larger journal segments because the metadata
server spends less time flushing the journal }\label{fig:throughput-journal}
\end{figure*}

@@

% purpose of the journal
As shown in Figure~\ref{fig:journal} CephFS uses RADOS (1) as a metadata store
for all information about files including the hierarchical namespace and (2) as
a staging area for the journal of updates before they are applied to the
metadata store. (1) is essentially a cache that improves performance by serving
metadata from memory and (2) is the mechanisms for achieving fault tolerance.

@@

\section{Results}

@@

%\section{notes}
%Linking clients into our custom libcephfs
%
%Use namespace's recursive data structure to put policies on subtrees
%- consistency: eventual vs. strong, global vs. local
%  - e.g., BatchFS/DeltaFS: eventual, local
%  - e.g., POSIX: strong, global
%  - e.g., PLFS: no consistency
%- fault tolerance: global vs. local
%  - e.g., CephFS: global
%  - e.g., BatchFS/DeltaFS: local
%
%Experimental Setup
%- Ceph: 9 OSDs, 1 MDS, 2 kernel client
%- Workload limitations: blah
%
%Workload: creates
%
%Baseline: 200K creates in the same directory
%- throughput: degrades at 950s
%- CPU utilization: more at 950s
%- inode cache: eviction dominate
%- inodes +- to cache: eviction dominate
%- per-disk throughput: RADOS not bottleneck
%
%Experiment 1: Interference
%
%\subsection{Baseline}
%Experiment 0: creates in the same directory
%- setup: why we use caching, we use the kernel client, how we circumvent max fragment size
%
%Experiment 0: creates with a stat
%- Hypothesis: metadata read pauses creates and requires a snapshot in time
%  - what is more of an overhead: pausing creates and getting a consistent view OR sucking up resources as it reads from RADOS?
%- can we delay snapshot?
%
%Experiment 1: creates with a readdir
%- Hypothesis: shows the cost of synchronization because on a write, the first client drops his caps
%- client0: create 100k, client1: stat at 2 mins
%
%Experiment 2: scale the number of files
%- See if the open/close spike occurs 
%- Try to see why open/close spike is allowed to happen
%- Try to disable all caching -- metadata writes don't ever re-use the inode -- we never ask for it again!
%- client0: create 100k, client1: touch at 2 mins
%
%Experiment 3: see how fast the cache satisfies a read
%- client0: create 100k, stat inodes
%- client0: create 100k, client1: stat inodes
%
%lient 0: creates, client 1 create(s)
